---
title: "ANLY 502 Final Project Report"
author: ""
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Title: 1994 Adult Census Income Dataset Analysis

# Project Topic:
The data set contains the 1994 census data with 14 variables and over 48000 records. The data extraction was done by Barry Becker in 1994 and donated by Ronny Kohavi and Barry Baker. The variable ranges from individuals’ education level, marital status, race etc., all of which may or may not have any effect on individuals’ income. There has been some data manipulation by the provider of the data set. The file format is text, comma delimited.

# Team Member: Sokpheanal Huynh

# Dataset Description:
The data set were divided into two files; adult.data - the training dataset and adult.test - a testing dataset.
The total number of records combined from the two datasets is 48,842 rows.

## Link to the dataset:https://archive.ics.uci.edu/ml/datasets/Adult

## The dataset has 14 explanatory variables and 1 target variable:
1. age: age of the individual - continuous.
2. workclass: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked - categorical
3. fnlwgt: a weight value the Census Bureau used - continuous.
4. education: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool - categorical
5. education-num: number of years of education - continuous.
6. marital-status: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse - categorical
7. occupation: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces - categorical
8. relationship: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried - categorical
9. race: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black - categorical
10. sex: Female, Male - categorical
11. capital-gain: amount of capital gain - continuous
12. capital-loss: amount of capital loss - continuous
13. hours-per-week: number of hours worked a week - continuous.
14. native-country: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands - categorical
15. income: >50K, <=50k - categorical

# Question:

H0: People with a higher education degree (at least 14 years of education) earn above $50,000. 
HA: People with a higher education degree (at least 14 years of education) do not earn above $50,000.

#Tools:
R was used for the project. We employed R, and a number of its functions to load the data, perform data wrangling, and statistical data analysis.

# Load the data:
```{r load data}
#Reading the training data
adult.income.data.trained <- read.csv("adult.data.csv")
#Reading the testing data
adult.income.data.test <- read.csv("adult.test.csv",
                                   header = FALSE)
str(adult.income.data.test)
str(adult.income.data.trained)
#Adding column names to the testing dataset
colnames(adult.income.data.test) <- c("age",
                                      "workclass",
                                      "fnlwgt",
                                      "education",
                                      "education_num",
                                      "marital_status",
                                      "occupation",
                                      "relationship",
                                      "race",
                                      "sex",
                                      "capital_gain",
                                      "capital_loss",
                                      "hours_per_week",
                                      "native_country",
                                      "income")
#Adding column names to the training dataset
colnames(adult.income.data.trained) <- c("age",
                                         "workclass",
                                         "fnlwgt",
                                         "education",
                                         "education_num",
                                         "marital_status",
                                         "occupation",
                                         "relationship",
                                         "race",
                                         "sex",
                                         "capital_gain",
                                         "capital_loss",
                                         "hours_per_week",
                                         "native_country",
                                         "income")
#Combining the two datasets
adult_data <- rbind(adult.income.data.test, adult.income.data.trained)
summary(adult_data)
```
# Load the necessary packages
```{r load packages}
library(dplyr)
library(sqldf)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(stringr)
library(DATA606)
```

# Data Analysis

## Exploring the workclass explanatory variable
###Manipulate the income variable
```{r income data manipulation}
#Removing leading spaces in workclass column
adult_data$workclass <- str_squish(adult_data$workclass)
#Removing spaces and period for the target variable data string
adult_data$income <- ifelse(adult_data$income == ' <=50K.',
                            '<=50k',
                            as.character(adult_data$income))
adult_data$income <- ifelse(adult_data$income == ' >50K.', 
                            '>50k', 
                            as.character(adult_data$income))
#Categorizing the target variable: >50k is 1, and <= 50k is 0
adult_data$income <- ifelse(adult_data$income == '>50k', 
                            1, 
                            0)
adult_data$income
```
###Manipulate the workclass explanatory variable
```{r workclass data manipulation}
#Replacing ? fields with Unknowns for the workclass variable
adult_data$workclass <- ifelse(adult_data$workclass == '?', 
                               'Unknown', 
                               as.character(adult_data$workclass))
adult_data$workclass
#Counting numbers of people with salary >50k for each workclass
work_class <- sqldf('SELECT workclass,
                    count(workclass) as Count,
                    sum(income) as Above 
                    from adult_data
                    group by workclass') 
work_class
#Calculate the proportion of individuals with >50k in each workclass, then output in a table 
workclass_table <- data.frame(Class = work_class$workclass,
                              Proportion = work_class$Above / work_class$Count)
workclass_table
#Calculate the proportion of individuals with <=50k in each workclass
work_class$Below <- work_class$Count - work_class$Above
#Dropping the count column since we only need the above and below stats
work_class <- work_class[ ,c(1, 3, 4)]
#Summarizing the Above/Below statistics for each of the workclass
workclass <- melt(work_class,
                  id.vars = 'workclass')
workclass
```
###Creating plot for workclass
```{r create workclass plots}
#Creating the initial workclass plot
work_class_plot <- ggplot(workclass,
                          aes(x = workclass,
                              y = value,
                              fill = variable))
work_class_gg <-  work_class_plot + 
                  geom_bar(stat = 'identity') +
                  theme_bw() +
                  scale_fill_manual(values = c("green", "red")) +
                  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                  ggtitle('Proportions of above-paid within different classes')
#Calculating the proportion of people earning above 50k based on workclass, then visualize the numbers in a table
work_class_proportion <- tableGrob(t(1:16),
                                  rows = NULL)
grid.arrange(work_class_proportion, 
             work_class_gg,
             nrow = 2,
             as.table = TRUE,
             heights = c(1, 4))
```
###Workclass analysis conclusion: 
Most people work in the private sector. The private sector is also where the largest number of the population that earn more than 50k per year. However, in terms of the proportion, the self-employed group has the highest ratio of people earning more than 50k per year.

The private sector, which has the most people work in, has the largest number of population that earn more than 50K per year. However, in terms of the proportion, the self-employed people are the winner.

## Exploring the education explanatory variable
###Analyzing the education variable
```{r education}
#Querying numbers of people with salary >50k for each education level
education_var <- sqldf('SELECT education, count(education) as Count,
                        sum(income) as Above 
                        from adult_data 
                        group by education')
#Calculate the number of people earning below 50k for each education level 
education_var$Below <- education_var$Count - education_var$Above
#Calculate the proportion of individuals with <=50k in each education level
table <- data.frame(Class = education_var$education,
                    Proportion = education_var$Above / education_var$Count)
#Dropping the count column since we only need the above and below stats
education_var <- education_var[ ,c(1,3,4)]
#Summarizing the Above/Below statistics for each of the education level
edu <- melt(education_var,
            id.vars = 'education')
```

###Creating plot for education
```{r create education plots}
#Creating the initial education plot
edu_gg <- ggplot(edu,
                 aes(x = education,
                     y = value,
                     fill = variable)) +
            geom_bar(stat = 'identity',
                 position = 'stack') +
            theme_bw() +
            scale_fill_manual(values = c('green','red')) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
            ggtitle('Proportions of above-paid within different education level')
#Calculating the proportion of people earning above 50k based on education level, then visualize the numbers in a table
edu_tbl <- tableGrob(t(table),
                     rows = NULL)
grid.arrange(edu_tbl, 
             edu_gg,
             nrow = 2,
             as.table = TRUE,
             heights = c(1, 4))
```
###Education variable analysis conclusion: 
Most people has a high school diploma. However, in terms of the proportion, the Professional school group has the highest ratio of people earning more than 50k per year, with Doctorate coming in as a close second, then Masters group at third.

## Exploring the race explanatory variable

```{r race}
#Querying numbers of people with salary >50k for each race
race_var <- sqldf('SELECT race, count(race) as Count,
                   sum(income) as Above 
                   from adult_data 
                   group by race')
#Calculate the number of people earining below 50k for each race
race_var$Below <- race_var$Count - race_var$Above
#Calculate the proportion of individuals with <=50k in each race
race_table <- data.frame(race_var = race_var$race,
                         Proportion = race_var$Above / race_var$Count)
#Dropping the count column since we only need the above and below stats
race <- race_var[ ,c(1,3,4)]
#Summarizing the Above/Below statistics for each of the race
rac <- melt(race,id.vars = 'race')
```

###Creating plot for race
```{r create race plots}
#Creating the initial race plot
race_gg <- ggplot(rac,
                  aes(x = race,
                      y = value,
                      fill = variable)) +
                  geom_bar(stat = 'identity',
                           position = 'stack') +
                  theme_bw() +
                  scale_fill_manual(values = c('green', 'red')) +
                  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                  ggtitle('Proportions of above-paid within different races')
#Calculating the proportion of people earning above 50k based on race, then visualize the numbers in a table
race_tbl <- tableGrob(t(table),
                      rows = NULL)
grid.arrange(race_tbl,
             race_gg,
             nrow = 2,
             as.table=TRUE,
             heights = c(1, 4))
```
###Race variable analysis conclusion: 
White is the most represented race in the dataset. In addition, statistically only, they seem to be advantageous in salary comparing to other races.

## Exploring the sex variable
```{r sex}
#Querying numbers of people with salary >50k for each sex
sex_var <- sqldf('SELECT sex, 
                  count(sex) as Count,
                  sum(income) as Above 
                  from adult_data 
                  group by sex')
#Calculate the number of people earining below 50k for each sex
sex_var$Below <- sex_var$Count - sex_var$Above
#Calculate the proportion of individuals with <=50k in each sex
sex_table <- data.frame(sex = sex_var$sex, 
                        Proportion = sex_var$Above / sex_var$Count)
#Dropping the count column since we only need the above and below stats
sex <- sex_var[,c(1,3,4)]
#Summarizing the Above/Below statistics for each of the sex
se <- melt(sex,
           id.vars = 'sex')
se
```
###Creating plot for sex
```{r create sex plots}
#Creating the initial sex plot
sex_gg <- ggplot(se,
                 aes(x = sex,
                     y = value,
                     fill = variable)) + 
          geom_bar(stat = 'identity',
                   position = 'stack') +
          theme_bw() +
          scale_fill_manual(values = c('green', 'red')) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          ggtitle('Proportions of above-paid within different sexes')
#Calculating the proportion of people earning above 50k based on sex, then visualize the numbers in a table
sex_tbl <- tableGrob(t(table),
                     rows = NULL)
grid.arrange(sex_tbl,
             sex_gg,
             nrow = 2,
             as.table = TRUE,
             heights = c(1, 4))
```
###Sex variable analysis conclusion: 
The plot shows that there are significantly more male individuals sampled. In addition, male employees were earning more competitive salary than female employees.

## Explore Continuous Variables

## Explore the Capital Loss variable
```{r capital loss column}
#Renaming the capital loss column by removing the "_"
colnames(adult_data)[12] <- 'CapitalLoss'
```
###Creating the capital loss plot
```{r capital loss plot}
capital_loss_gg <- qplot(CapitalLoss,
                         data = adult_data,
                         geom = "histogram") +
                         theme_bw() +
                         ggtitle('Histogram of Capital Loss')
capital_loss_gg
```
###Capital Loss variable analysis connclusion: 
The capital loss variable is quite skewed and mostly concentrated at aroudn the zero value. As such, it may not be useful for classification.

## Explore the Working hours variable
```{r working hours data}
#Updating the column name from hours_per_week to Hours
colnames(adult_data)[13] <- 'Hours'
```
```{r working hours preliminary plot}
#Creating a histogram for the working hours variable
hours_gg <- qplot(Hours,
                  data = adult_data,
                  geom = "histogram") +
                  theme_bw() +
                  ggtitle('Histogram of Working Hours')

hours_gg
```
```{r catgorizing working hours}
#Reload the data for further data manipulation
adult_data2 <- adult_data
#Categorizing the working hours to NormalWorkload (<= 40 hours) and HugeWorkLoad (> 40 hours)
adult_data2$HourJ <- ifelse(adult_data2$Hours <= 40,
                            'NormalWorkLoad',
                            'HugeWorkLoad')
#Querying the count of people with huge work load vs normal workload, and the number of people earning > 50k in each category
work_load <- sqldf('SELECT HourJ as WorkLoad,
                    count(HourJ) as Count,
                    sum(income) as Above 
                    from adult_data2 
                    group by HourJ')
#Calculate the number of people earning below 50k for each work load
work_load$Below <- work_load$Count - work_load$Above
#Calculate the workload percentage
work_load_percentage <- work_load$Above / work_load$Count
#Dropping the count column since we only need the above and below stats
work_load <- work_load[ ,c(1,3,4)]
#Summarizing the Above/Below statistics for each of the sex
wlt <- melt(work_load,
            id.vars = 'WorkLoad')
wlt
#Adding the workload percentage column to the work_load table
wl <- cbind(work_load,
            work_load_percentage)
```
###Creating plots for the workload variable
```{r workload plots}
#Create preliminary plot for workload
work_load_gg <- ggplot(wlt,
                       aes(x = WorkLoad,
                           y = value,
                           fill = variable)) +
                geom_bar(stat = 'identity',
                          position = 'stack') +
                theme_bw() + 
                scale_fill_manual(values = c('green', 'red')) +
                theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
                ggtitle('Proportions of above-paid with different Work Load')
#Calculating the proportion of people earning above 50k based on work load, then visualize the numbers in a table
work_load_tbl <- tableGrob(t(wl[,c(1,4)]), 
                           rows = NULL)
grid.arrange(work_load_tbl, 
             work_load_gg,
             nrow = 2,
             as.table = TRUE,
             heights = c(1, 4))
```
###Work load variable analysis conclusion: 
There is a higher proportion of people with bigger workload who earn more than 50k.

# Hypothesis testing:

H0: People with a higher education degree (at least 14 years of education) earn above $50,000. 
HA: People with a higher education degree (at least 14 years of education) do not earn above $50,000.

##Calculate a 95% confidence interval for higher education (more than 14 years of education)
We first need to copy the original dataset to a new one to perform a separate analysis. In addition, the education_num variable would be categorized, >= 14 years of education and < 14 years of education. 
```{r data manipulation}
#Manipulate the education_num variable
#Copying data from the original data to perform separate analysis.
adult_data3 <- adult_data
#Categorizing the education_num variable; If >= 14 years then show >= 14, else < 14
adult_data3$education_num <- ifelse(adult_data3$education_num >= 14,
                                    ">= 14",
                                    "< 14")
#Manipulate the income variable
#Update the income variable; change the data back to show >= 50 when 1, and <50 when 0
adult_data3$income <- ifelse(adult_data3$income == 1, 
                            ">= 50k", 
                            "< 50k")
#Show the statistics of income level vs number of years of education categories in a table
table(adult_data3$education_num,
      adult_data3$income)
```
Base on the table that shows the statistics of income level (< 50k or >= 50k) versus the number of years of education (< 14 years or >= 14 years), we can see thata majority of people were earning less than 50k and have less than 14 years of education.

We then set up our experiment by dividing the data set, using the subset function, into 2 groups, a group with earnings at least 50k (>= 50k) and a group with earnings less than 50k.
```{r further manipulation}
#Set up experiment by subsetting the dataset to a group of earnings at least 50k and a group of earnings below 50k
at_least_50k <- subset(adult_data3, 
                      income == ">= 50k")
str(at_least_50k)
below_50k <- subset(adult_data3,
                   income == "< 50k")
str(below_50k)
```
#Inference for the >= 50k and >= 14 years of education group
```{r inference at least 50k with at least 14 years}
inference(at_least_50k$education_num,
          est = "proportion",
          type = "ci",
          method = "theoretical",
          success = ">= 14")
#Calculating proportion of people earning at least 50k
prop_min_50k_atleast14 <- 819/3846
prop_min_50k_atleast14
```
##Inference on the >= 50k and >= 14 years of education group conclusion:
The confidence interval of the group with the years of education equal to or higher than 14 while earning more than 50k is ( 0.2 , 0.2259 ).
The P-value of the group that earns at least 50k while having at least 14 years of education is 0.2129. 
The proportion of people earning more than 50k in 1994 with at least 14 years of education is 0.25077.
With the p-value found well above the 0.05 or 5% threshold, the p-value is not statistically significant. Hence, the null hypothesis, which is "People with a higher education degree (at least 14 years of education) earn above 50k", is rejected, and the alternate hypothesis; that is, people with a higher education degree (at least 14 years of education) do not earn above 50k, is accepted.

#Further question
Now that the null hypothesis is rejected, we would like to examine if the education_num is, in any case, a good predictor of the income level by using the inference function on the group with income less than 50K while having less than 14 years of education.

##Inference for the < 50k and < 14 years group
```{r inference for the less than 50k and less than 14 years}
inference(below_50k$education_num,
          est = "proportion",
          type = "ci",
          method = "theoretical",
          success = "< 14")
#Calculating proportion of people earning at least 50k while having less than 14 years of education
prop_below_50k_lessthan14 <- 41730/44996
prop_below_50k_lessthan14
```
##Inference on the < 50k and < 14 years of education group conclusion:
The confidence interval of the group with the years of education less than 14 while earning less than 50k is ( 0.925 , 0.9298 ).
The P-value of the group that earns less than 50k while less than 14 years of education is 0.9274. 
The proportion of people earning more than 50k in 1994 with at least 14 years of education is 0.9274158.
With the p-value found well above the 0.05 or 5% threshold, the p-value is not statistically significant.
In either cases, the p-value were not statistically significant. As a result, we can conclude that the number of years of education is not an accurate predictor of the income level. 

#Findings:
Through the analysis, we have come to the findings below:
1. The private sector, which has the most people work in, has the largest number of population that earn more than 50K per year. However, in terms of the proportion, the self-employed people are the winner.
2. Most people has a highschool diploma. However, in terms of the proportion, the Professional school group has the highest ratio of people earning more than 50k per year, with Doctorate coming in as a close second, then Masters group at third.
3. White is the most represented race in the dataset. In addition, statistically only, they seem to be advantageous in salary comparing to other races.
4.The plot for the sex data analysis shows that there are significantly more male individuals sampled. In addition, male employees were earning more competitive salary than female employees.
5. The capital loss variable is quite skewed and mostly concentrated at aroudn the zero value. As such, it may not be useful for classification.
6. There is a higher proportion of people with bigger workload who earn more than 50k.
7. People with a higher education degree (at least 14 years of education) do not earn above $50,000.
8. The number of years of education is not an accurate predictor of the income level.

#Conclusion
Through the analysis of this dataset, we were able to come to numerous interesting findings. The null hypothesis rejection shows that there is no real correlation between an individual's level of education and the income level, which goes against the popular beliefs. Further analysis can be done to determine the relationships between the explanatory variables and the target income variable; for instance, the combination of age and education_num vs income. In addition, we have found that entrepreneurs group are more likely to earn more than 50K, comparing to other groups. The dataset also represented the earning gaps between males and females, which is, currently, still an issue in our society/