---
title: "Aggregating and Formatting Data"
author: "Sokpheanal Huynh"
date: "rev 2023-04-10"
output: html_document
---

# Instructions

In class we reviewed methods for collecting data from more than one source.

Please perform the following:

*   [X] Display the aggregated data as a table.

# Results
The Dallas Cowboys had the highest number of wins in the 1971-1972 season.

# Code

```{r load files}
games_hw <- read.csv("Aggregating and Formatting Data (1).csv")
teams_hw <- read.csv("Aggregating and Formatting Data (2).csv")
str(teams_hw)
#teams_hw
str(games_hw)
#games_hw
```
```{r combine the files, message = F, warning = F}
library(dplyr)
(t1_hw <-
  games_hw %>%
  select(id = Winner.tie))

#Winning team table
(t1_hw <-
   t1_hw %>%
   left_join(teams_hw, by = c('id' = 'id')))
```

```{r Calculate the number of wins per team}
(total_win_by_team <-
  t1_hw %>%
  group_by(teams) %>%
    summarise(
      total_win = n()))
```
```{r rank the teams}
(total_win_by_team_ranked <-
   total_win_by_team %>%
   mutate(rank = dense_rank(desc(total_win))) %>%
   arrange(rank))

#rearrange the columns position
(total_win_by_team_ranked <-
    total_win_by_team_ranked %>%
    relocate(rank, .before = total_win))

#Updating column names to match prompt
colnames(total_win_by_team_ranked) <- c('Teams', 'Rank', 'Total Wins')
```
```{r write data to csv}
write.csv(total_win_by_team_ranked, "C:/Users/sokpheanal.huynh/OneDrive - Envestnet/ANLY560/AggregatingAndFormattingHomeworkResults.csv", row.names=TRUE)
```


# Tasks
Load your data and create a knitted() table.

* Display the team and their rank

# Steps

* Load both files
* Combine the files
* Calculate the number of wins per team
* Rank the teams
* Save the results

# Results
Load the data from your saved file. Display the results as in the example below, with the best team on top.

```{r install and library kable, echo = F, message = F}
library(kableExtra)
library(knitr)
```
```{r table, echo = F}
(total_win_by_team_ranked %>%
   kbl() %>%
   row_spec(0, align = 'c') %>%
   kable_styling())
```