---
title: "Prediction using Association"
author: "Sokpheanal Huynh"
date: "2023/06/06"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Outline

* [X] Examine data
* [X] Formulate questions
* [X] Plan analysis - Pseudo Code
* [X] Plan charts
* [X] Create Predictions
* [X] Create Chart

# Data

"Real" data can be found [here](https://www.pro-football-reference.com/years/1971/games.htm) and [here](https://en.wikipedia.org/wiki/National_Football_League_franchise_moves_and_mergers).

# Solution

````{r}
data_1971 <- read.csv('1971.csv')
data_2021 <- read.csv('2021.csv')

str(data_1971)
str(data_2021)
View(data_1971)
View(data_2021)
```

QC

```{r}
ncol(data_1971) == ncol(data_2021)
all(colnames(data_1971) == colnames(data_2021))

str(data_1971)
class(data_1971)
class(data_1971$Week)

all(sapply(data_1971, class) == sapply(data_2021, class))
```

```{r}
game_files <- list.files(pattern = '[0-9]{4}\\.csv')

library(lubridate)
library(dplyr)

games_data_raw <- vector(mode = 'list', length = length(game_files))
for(i in 1:length(games_data_raw)) {
  games_data_raw[[i]] <- 
    read.csv(game_files[i]) %>%
    filter(Date != 'Playoffs') %>%
    mutate(Date = ymd(Date),
           Season = min(year(Date)))
   

  qc1 <- print(ncol(games_data_raw[[1]]) == ncol(games_data_raw[[i]]))
  qc2 <- all(colnames(games_data_raw[[1]]) == colnames(games_data_raw[[i]]))
  qc3 <- all(sapply(games_data_raw[[1]], class) == sapply(games_data_raw[[i]], class))

  if(all(qc1, qc2, qc3)) {} else { stop('QC failed') }
}
games_data <- do.call(rbind, games_data_raw)
```

```{r}
crosswalk_file <-'crosswalk.csv'
crosswalk_data <- read.csv(crosswalk_file)

str(crosswalk_data)
View(crosswalk_data)
```

```{r}
#install.packages('tidyr')
library(tidyr)

game_data <-
  games_data %>%
  left_join(crosswalk_data, by = c('Winner.tie' = 'Old')) %>%
  mutate(Winner.tie = ifelse(is.na(New), Winner.tie, New), New = NULL) %>%
  left_join(crosswalk_data, by = c('Loser.tie' = 'Old')) %>%
  mutate(Loser.tie = ifelse(is.na(New), Loser.tie, New))
```

```{r}
rm(crosswalk_data, games_data_raw, crosswalk_file, game_files, i, qc1, qc2, qc3, data_1971, data_2021, games_data)
```

# Questions

Given a specific game, can we tell who is more likely to win?

* Odds Ratios
* History
  * How much? 10
  * Same teams - no
  * home vs away - no

# Steps

1. Filtering
2. Odds Ratio ~ calculate the number of wins and losses for history(10) per game
3. Calculate if we are right
4. Roll-up accuracy to a season level
5. Charts - top-line accuracy

```{r}
game_data2 <-
  game_data %>%
  filter(Pts != Pts.1) %>%
  select(
    Season,
    Date,
    Time,
    Winner = Winner.tie,
    Loser = Loser.tie)
```

Predict (by hand)

* Denver 2 and 8
* San Diego 4 and6

```{r}
game_data[340,]
game_data2 <- game_data
```

Predict (little code)

```{r}
n <- 10
team <- 'Denver Broncos'

t1 <-
  game_data %>%
  filter(Date < ymd('1972-12-10')) %>%
  filter(Winner == team | Loser == team) %>%
  arrange(desc(Date)) %>%
  slice_head(n = n)
sum(t1$Winner == team)
sum(t1$Loser == team)
```


```{r}
team <- 'San Diego Chargers'

t1 <-
  game_data %>%
  filter(Date < ymd('1972-12-10')) %>%
  filter(Winner == team | Loser == team) %>%
  arrange(desc(Date)) %>%
  slice_head(n = n)
sum(t1$Winner == team)
sum(t1$Loser == team)
```

* Denver Broncos 2 and 8
* San Diego Chargers 4 and 6

Predict (rolling wins - single team)

```{r}
n <- 10
team <- 'Denver Broncos'

t1 <-
  game_data %>%
  filter(Date <= ymd('1973-12-10')) %>%
  filter(Winner == team | Loser == team) 

t2 <-
  t1 %>%
  transmute(
    Season,
    Date,
    Team = Winner,
    Win = 1)
t3 <-
  t1 %>%
  transmute(
    Season,
    Date,
    Team = Loser,
    Win = 0)
t4 <-
  rbind(t2, t3) %>%
  filter(Team == team) %>%
  arrange(Date)
```

Roll

```{r}
library(RcppRoll)

t5 <-
  t4 %>%
  mutate(
    WinR = roll_sum(Win, n = n, align = 'right', fill = NA),
    WinRS = c(NA, WinR[1:(n()-1)]),
    LossR = roll_sum(!Win, n = n, align = 'right', fill = NA),
    LossRS = c(NA, LossR[1:(n()-1)]))
t5[26,c('WinRS', 'LossRS')]
```

```{r}
n <- 10
team <- 'San Diego Chargers'

t1 <-
  game_data %>%
  filter(Date <= ymd('1973-12-10')) %>%
  filter(Winner == team | Loser == team) 
t2 <-
  t1 %>%
  transmute(
    Season,
    Date,
    Team = Winner,
    Win = 1)
t3 <-
  t1 %>%
  transmute(
    Season,
    Date,
    Team = Loser,
    Win = 0)
t4 <-
  rbind(t2, t3) %>%
  filter(Team == team) %>%
  arrange(Date) %>%
  mutate(
    WinR = roll_sum(Win, n = n, align = 'right', fill = NA),
    WinRS = c(NA, WinR[1:(n()-1)]),
    LossR = roll_sum(!Win, n = n, align = 'right', fill = NA),
    LossRS = c(NA, LossR[1:(n()-1)]))
t4[nrow(t4), c('WinRS', 'LossRS')]  
```

* Denver Broncos 2 and 8
* San Diego Chargers 4 and 6

Predict (Rolling wins - all teams/all dates)

```{r}
n <- 10
t2 <-
  game_data %>%
  transmute(
    Season,
    Date,
    Team = Winner,
    Win = 1)
t3 <-
  game_data %>%
  transmute(
    Season,
    Date,
    Team = Loser,
    Win = 0)
rolling_odds <-
  rbind(t2, t3) %>%
  arrange(Date) %>%
  group_by(Team) %>%
  mutate(
    WinR = roll_sum(Win, n = n, align = 'right', fill = NA),
    WinRS = c(NA, WinR[1:(n()-1)]),
    LossR = roll_sum(!Win, n = n, align = 'right', fill = NA),
    LossRS = c(NA, LossR[1:(n()-1)])) %>%
  select(Season, Date, Team, WinRS, LossRS)
```

QC

```{r}
rolling_odds %>%
  filter(Date == ymd('1972-12-10')) %>%
  filter(Team == 'Denver Broncos' | Team == 'San Diego Chargers')

game_data %>%
  filter(Date == ymd('1972-12-10')) %>%
  filter(Winner == 'Denver Broncos' | Winner == 'San Diego Chargers')
```

Predict (full)

* DF Team 1 and Team 2 for each game
* DF needs to not leak info
* Use 'rolling_odds' to predict result
* Compare that to ground truth ('game_data')

```{r}
set.seed(0)

game_data <- 
  game_data %>%
  mutate(ID = row_number())

ground_truth <-
  game_data %>%
  select(ID, Winner)

indx <- sample(1:2, size = nrow(game_data), replace= T)
games <-
  game_data %>%
  mutate(
    Team1 = ifelse(indx == 1, Winner, Loser),
    Team2 = ifelse(indx == 2, Winner, Loser)) %>%
  select(ID, Season, Date, Team1, Team2)
```

Join

```{r}
t1 <-
  games %>%
  left_join(rolling_odds, by = c('Season' = 'Season', 'Date' = 'Date', 'Team1' = 'Team')) %>%
  mutate(
    Team1_Odds = WinRS/LossRS,
    WinRS = NULL,
    LossRS = NULL) %>%
  left_join(rolling_odds, by = c('Season' = 'Season', 'Date' = 'Date', 'Team2' = 'Team')) %>%
  mutate(
    Team2_Odds = WinRS/LossRS,
    WinRS = NULL,
    LossRS = NULL)
```

QC

```{r}
nrow(t1) == nrow(games)
```

Make Prediction

```{r}
prediction <-
  t1 %>%
  mutate(
    Winner = ifelse(Team1_Odds > Team2_Odds, Team1, Team2)) %>%
  select(ID, Winner)
```

Compare

```{r}
results <-
  ground_truth %>%
  left_join(prediction, by = c('ID' = 'ID')) %>%
  select(
    ID,
    Actual = Winner.x,
    Predicted = Winner.y) %>%
  mutate(
    Correct = Actual == Predicted)

sum(results$Correct, na.rm = T)
sum(!results$Correct, na.rm = T)
```

Results: 60% overall

```{r}
t1 <-
  games %>%
  select(ID, Season)
t2 <-
  results %>%
  left_join(t1, by = c('ID' = 'ID'))

t3 <-
  t2 %>%
  group_by(Season) %>%
  summarise(
    Wrong = sum(!Correct, na.rm = T),
    Correct = sum(Correct, na.rm = T),
    Total = Correct + Wrong,
    N = n(),
    ACC = Correct/Total)
```

```{r}
library(ggplot2)

ggplot(t3, aes(x = Season, y = ACC)) +
  theme_bw() +
  geom_line() + 
  geom_hline(yintercept =  0.5)
```



